Precision calculations of the cosmic shear power spectrum projection

### 2017-02-16 authors.tex
Removed line about flat universe in appendix

\author[M.~Kilbinger et~al.]
{Martin Kilbinger$^{1,2}$\thanks{E-mail: martin.kilbinger@cea.fr},
Catherine Heymans$^3$,
Marika Asgari$^3$, 
Shahab Joudaki$^{4,5}$, 
\newauthor
Peter Schneider$^6$,
Patrick Simon$^6$,
Ludovic Van Waerbeke$^8$,
Joachim Harnois-D\'eraps$^3$,
\newauthor
Hendrik Hildebrandt$^6$,
Fabian K\"ohlinger$^7$,
Konrad Kuijken$^7$, and
Massimo Viola$^7$
\\
$^1$CEA/Irfu/SAp Saclay, Laboratoire AIM, 91191 Gif-sur-Yvette, France\\
$^2$Institut d'Astrophysique de Paris, UMR7095 CNRS, Universit\'e Pierre \& Marie Curie, 98 bis boulevard Arago, 75014 Paris, France \\
$^3$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK\\
$^4$Centre for Astrophysics \& Supercomputing, Swinburne University of Technology, PO Box 218, Hawthorn, VIC 3122, Australia\\
$^5$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
$^6$Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany\\
$^7$Kavli Institute for the Physics and Mathematics of the Universe (WPI), The University of Tokyo Institutes for Advanced Study, \\
The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^8$Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada\\
}

>>>>

\author[M.~Kilbinger et~al.]
{Martin Kilbinger$^{1,2}$\thanks{E-mail: martin.kilbinger@cea.fr},
Catherine Heymans$^3$,
Marika Asgari$^3$, 
Shahab Joudaki$^{4,5}$, 
\newauthor
Peter Schneider$^6$,
Patrick Simon$^6$,
Ludovic Van Waerbeke$^8$,
Joachim Harnois-D\'eraps$^3$,
\newauthor
Hendrik Hildebrandt$^6$,
Fabian K\"ohlinger$^7$,
Konrad Kuijken$^7$, and
Massimo Viola$^7$
\\
$^1$CEA/Irfu/SAp Saclay, Laboratoire AIM, 91191 Gif-sur-Yvette, France\\
$^2$Institut d'Astrophysique de Paris, UMR7095 CNRS, Universit\'e Pierre \& Marie Curie, 98 bis boulevard Arago, 75014 Paris, France \\
$^3$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK\\
$^4$Centre for Astrophysics \& Supercomputing, Swinburne University of Technology, PO Box 218, Hawthorn, VIC 3122, Australia\\
$^5$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
$^6$Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany\\
$^7$Kavli Institute for the Physics and Mathematics of the Universe (WPI), The University of Tokyo Institutes for Advanced Study, \\
\phantom{$^7$}The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^8$Department of Physics and Astronomy, University of British Columbia, 6224 Agricultural Road, Vancouver, BC V6T 1Z1, Canada\\
}

### 2017-02-16 cfhtlens.tex
Changes from Marika added
Removed line about flat universe in appendix

analytical methods will be published soon (Joachimi et al.~in prep.}

@@

For a first-order standard Limber flat-sky approximation (L1Fl) we find
$\sigma_8 (\Omega_{\rm m}/0.27)^{0.6} =0.787^{+0.031}_{-0.033}$, the same
result that was published in \cite{CFHTLenS-2pt-notomo}. Using the second-order
extended Limber flat-sky hybrid approximation (ExtL2FlHyb) results in $\sigma_8
(\Omega_{\rm m}/0.27)^{0.6} = 0.788 \pm 0.032$, a negligible change of the
amplitude that is well within the Monte-Carlo sampling noise. The largest
difference is measured with the depreciated case ExtL1Fl, for which the
recovered amplitude is larger by $16\%$ of the statistical error. \ch{These
negligible changes to the error bars were to be expected owing to the
high level of statistical noise and cosmic variance in comparison to the low-level impact of the various
approximations shown in Fig.~\ref{fig:Cl_cases}.}

@@

\ch{Considering the cosmological constraints from tomographic Kilo-Degree Survey (KiDS), we conclude that
these are robust to flat-sky and Limber approximations. The case ExtL1FlHyb that was used
for the analysis of KiDS data in \citet{KiDS-450} and \cite{joudaki/etal:2017} introduces
errors that are more than an order of magnitude lower than the cosmic variance
for that survey, and thus this approximation has a negligible impact on the
cosmological parameters.}

>>>>

analytical methods will be published soon (Joachimi et al.~in prep.)}

@@

For a first-order standard Limber flat-sky approximation (L1Fl) we find
$\sigma_8 (\Omega_{\rm m}/0.27)^{0.6} =0.787^{+0.031}_{-0.033}$, the same
result that was published in \cite{CFHTLenS-2pt-notomo}. Using the second-order
extended Limber flat-sky hybrid approximation (ExtL2FlHyb) results in $\sigma_8
(\Omega_{\rm m}/0.27)^{0.6} = 0.788 \pm 0.032$, a negligible change of the
amplitude that is well within the Monte-Carlo sampling noise. The largest
difference is measured with the depreciated case ExtL1Fl, for which the
recovered amplitude is larger by $16\%$ of the statistical error. These
negligible changes to the error bars were to be expected owing to the high
level of statistical noise and cosmic variance in comparison to the low-level
impact of the various approximations shown in Fig.~\ref{fig:Cl_cases}.

@@

Considering the cosmological constraints from tomographic Kilo-Degree Survey
(KiDS), we conclude that these are robust to flat-sky and Limber
approximations. The case ExtL1FlHyb that was used for the analysis of KiDS data
in \citet{KiDS-450} and \cite{joudaki/etal:2017} introduces errors that are
more than an order of magnitude lower than the cosmic variance for that survey,
and thus this approximation has a negligible impact on the cosmological
parameters.

### 2017-02-16 limber2.tex
Changes from Marika added
Removed line about flat universe in appendix

For future
large-scale surveys with unprecedented precision, we show that the spherical
second-order Limber approximation will provide sufficient accuracy. \ch{In this case, the
cosmic-shear power spectrum is shown to be in agreement with the full projection at the
sub-percent level for $\ell > 3$, with the corresponding errors an order of
magnitude below cosmic variance for all $\ell$.
When computing the two-point shear correlation function, we show that the flat-sky fast Hankel 
transformation results in errors below two percent compared to the full spherical transformation.}
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case equations~(\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations can be made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, \ch{DES-SV \citep{abbott/etal:2016} and DLS \citep{jee/etal:2016} analyses, 
    this was the approximation of choice.}
    Note that we do not discuss the second-order Limber approximation \ch{with $p(\ell)=1$}.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl, ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation~\LabelTxt{ExtL1FlHyb, ExtL2FlHyb}, and the first-order case 
    was used in \cite{KiDS-450,joudaki/etal:2016,joudaki/etal:2017}. As is shown
    below, this is a better approximation to the full projection than case (ii). In equation~(\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Considering first the flat-sky cases, the standard first-order Limber
approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, \ch{DES-SV and DLS analyses}, we
find it to be accurate to better than 10\% for $\ell>3$, converging slowly to the
true projection with percent level precison at $\ell>100$. For the extended
Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite
decreased accuracy for $\ell < 8$ in comparison to the standard first-order
Limber case, the errors with respect to the true power spectrum decrease much
faster, as $\ell^{-2}$, such that percent-level precision is reached at
$\ell>15$.  The first-order extended Limber approximation `hybrid' cases
(ExtL1FlHyb) was adopted by \citet{joudaki/etal:2016}, \citet{joudaki/etal:2017}, and
\cite{KiDS-450}\footnote{We confirm that there is a typographical error in
equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in
$\nu(\ell)$ that was incorporated in the cosmological analysis.}.

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  %
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. 'ID' is the label
    used in the text and figures.
    The third column shows the power-spectrum pre-factor $p(\ell)$. } 
    %
    \begin{tabular}{|l|l|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & Comment \\ \hline
  % 
  $1^{\rm st}$-order standard Limber, flat sky & L1Fl & (\ref{eq:P_ell_limber1}) + (i)
    & $1$ & $\ell$ & \ch{Pre-2014 CFHTLenS, DES-SV and DLS} \\ \hline
  %
  $1^{\rm st}$-order extended Limber, flat sky & ExtL1Fl & (\ref{eq:P_ell_limber1}) + (ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $1^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL1FlHyb & (\ref{eq:P_ell_limber1}) + (iii)
    & $1$ & $\ell + \frac 1 2$ & \ch{Post-2014 CFHTLenS, KiDS} \\ \hline
  %
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & \\ \hline
  %
  $2^{\rm st}$-order extended Limber, flat sky & ExtL2Fl &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $2^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL2FlHyb &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & Best flat-sky approximation \\ \hline
  %
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & Best approximation \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      - & - & Correct projection \\ \hline
  %
  \end{tabular}

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by the \mk{spherical} transform \citep{1973ApJ...185..413P}
%
\begin{equation}
 \xi_+(\theta) = \mk{\left\langle \gamma \gamma^\ast \right\rangle(\theta)}
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$. \mk{This equation is easily derived
using the harmonics expansion of the shear (\ref{eq:gamma_harm_exp}) and the orthogonality relation of
the spherical harmonics (\ref{eq:Yellm_ortho}).}

@@

We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to reduce computation time.
%
\ch{We find that the Hankel transform (equation~\ref{eqn:xiGG}) is accurate to better than $2$ percent level.
%
The difference between second-order Limber and full projection using the
spherical transform (\ref{eq:xi_Legendre}) is well below one percent
for $\vartheta < 200$ arcminutes.}
The red circles in Fig.~\ref{fig:xi_p_legendre} present the limit of
precision that we can achieve with the current fast Hankel transform implementation
for the correlation function. The green line shows the limit of the
second-order Limber approximation on the correlation function.

@@

  \caption{\emph{Left panel:} The shear correlation function component $\xi_+$
for the ExtL2Sph projection using the spherical transform (green line for ExtL2Sph, grey squares for FullSph),
compared to the flat-sky Hankel transform of ExtL2Sph (red circles).
\emph{Right panel:} The relative
difference \mk{compared to the full spherical case FullSph of ExtL2Sph with spherical transform
(green line), 
and Hankel transform (red circles).}}

@@

To obtain the power spectrum of the lensing potential for a flat Universe, 
we insert the lensing projection (equation~\ref{eq:psi}) for $K=0$ into the
inverse harmonics expansion (equation~\ref{eq:psi_harm_exp_inv}) and write the 3D potential
as its Fourier transform (equation~\ref{eq:hatPhi_inv}) to get
%
\begin{equation}
  \psi_{\ell m} = \frac 2 {c^2} \int {\rm d} \Omega {\rm Y}^\ast_{\ell m}(\theta, \varphi)
    \int_0^\infty \frac{{\rm d}\chi}{\chi} q(\chi) \int \frac{{\rm d}^3 k}{(2\pi)^3} \hat \Phi(\vec k; \chi) {\rm e}^{-{\rm i} \vec k \cdot \vec r}.
\end{equation}
%
The 3D position vector $\vec r$ is a 3D position vector with polar coordinate
$r = \chi$ and polar angles $(\theta, \varphi)$. Similarly we denote with
$\theta_k, \varphi_k$ the polar angles of the 3D Fourier vector $\vec k$. We
insert the expansion of a plane wave into spherical harmonics,
%
% BBRV12 (41)
%
\begin{equation}
  {\rm e}^{{\rm i} \vec k \cdot \vec r} = 4 \pi \sum_{\ell=0}^{\infty} \sum_{m=-\ell}^{\ell}
    {\rm i}^\ell \, {\rm j}_\ell(k \chi)
    {\rm Y}_{\ell m}(\theta, \varphi) {\rm Y}_{\ell m}(\theta_k, \varphi_k) .
  \label{eq:wave_exp}
\end{equation}
%
Making use of the orthogonality of the spherical harmonics
%
% Rijt (6.15), Castro (B2).
%
\begin{equation}
  \int {\rm d} \Omega {\rm Y}^{}_{\ell m}(\theta, \varphi) {\rm Y}^\ast_{\ell^\prime m^\prime}(\theta, \varphi) = \delta_{\ell \ell^\prime} \delta_{m m^\prime},
  \label{eq:Yellm_ortho}
\end{equation}
%
the expression for $\psi_{\ell m}$ simplifies to
%
\begin{equation}
  \psi_{\ell m} = \frac {{\rm i}^\ell}{c^2 \pi^2} \int_0^\infty \frac{{\rm d}\chi}{\chi} q(\chi) \int {\rm d}^3 k \,
    \hat \Phi(\vec k; \chi) {\rm j}_\ell(k \chi) {\rm Y}_{\ell m}(\theta_k, \varphi_k).
  \label{eq:psi_ellm}
\end{equation}
%
To obtain the potential power spectrum, we take the absolute square of the
last equation and use the definition of the 3D potential power spectrum
(equation~\ref{eq:p_phi}). The delta-function resolves one 3D Fourier integral.
We split the second integration into radial and spherical coordinates, ${\rm
d}^3 k = {\rm d} k k^2 {\rm d} \Omega_k$ and use once again the orthogonality
of the spherical harmonics to resolve the spherical integral. This leads to the
potential power spectrum in equation~(\ref{eq:C_ell_phi_Pphi}).

@@

\cite{2008PhRvD..78d3002S} derive the lensing power spectrum in the flat-sky
limit, see their Equation (9). Inserting the Poisson equation and the
growth function, and writing the redshift filter function
$W_\kappa$ (eq.~10) in terms of the lensing efficiency and comoving distances,
$W_\kappa[z(\chi)] = H(z)^{-1} \chi q(\chi)$, their expression, using our
notations, reads
%
\begin{align}
  P_{ij}^\gamma(\ell) = & \frac 2 \pi \, \pref^2
                 \int_{0}^\infty {\rm d} \chi \chi \, \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^\infty {\rm d} \chi^\prime\, \chi^\prime
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty {\rm d} k \, k^2 \, P_{\rm m}(k; \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) ,
  \label{eq:P_ell_gamma}
\end{align}
%
This is consistent with our equation (\ref{eq:P_ell_gamma_full_der}) under the
additional assumption that mainly modes with $k \chi \approx k \chi^\prime
\approx \ell$ contribute to the integral; that is modes around the maxima of
the Bessel functions. Then, we can draw out of the integral the factor $\ell^4
\approx k^4 \chi^2 {\chi^\prime}^2$, to recover
equation~(\ref{eq:P_ell_gamma_full_der}). This argument is very hand-wavey!
In fact, using the approximation $k \chi = \ell$ seems to go too far, this is
already halfway the Limber approximation.

>>>>

For future large-scale surveys with unprecedented precision, we show that the
spherical second-order Limber approximation will provide sufficient accuracy.
In this case, the cosmic-shear power spectrum is shown to be in agreement with
the full projection at the sub-percent level for $\ell > 3$, with the
corresponding errors an order of magnitude below cosmic variance for all
$\ell$. When computing the two-point shear correlation function, we show that
the flat-sky fast Hankel transformation results in errors below two percent
compared to the full spherical transformation.
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case equations~(\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations can be made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl} Until recently,
  i.e.~for all pre-2014 CFHTLenS results, DES-SV \citep{abbott/etal:2016} and DLS
  \citep{jee/etal:2016} analyses, this was the approximation of choice. Note that
  we do not discuss the second-order Limber approximation with $p(\ell)=1$.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the
  \emph{extended} Limber approximation \LabelTxt{ExtL1Fl, ExtL2Fl} with
  $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation~\LabelTxt{ExtL1FlHyb, ExtL2FlHyb}, and the first-order case 
    was used in \cite{KiDS-450,joudaki/etal:2016,joudaki/etal:2017}. As is shown
    below, this is a better approximation to the full projection than case (ii). In equation~(\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Considering first the flat-sky cases, the standard first-order Limber
approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, DES-SV and DLS analyses, we
find it to be accurate to better than 10\% for $\ell>3$, converging slowly to the
true projection with percent level precison at $\ell>100$. For the extended
Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite
decreased accuracy for $\ell < 8$ in comparison to the standard first-order
Limber case, the errors with respect to the true power spectrum decrease much
faster, as $\ell^{-2}$, such that percent-level precision is reached at
$\ell>15$.  The first-order extended Limber approximation `hybrid' cases
(ExtL1FlHyb) was adopted by \citet{joudaki/etal:2016}, \citet{joudaki/etal:2017}, and
\cite{KiDS-450}\footnote{We confirm that there is a typographical error in
equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in
$\nu(\ell)$ that was incorporated in the cosmological analysis.}.

@@

\renewcommand{\baselinestretch}{1.5}
\begin{table}
\begin{centering}
  %
  \caption{\label{tab:cases}The shear power spectrum approximations studied in this paper. 'ID' is the label
    used in the text and figures.
    The third column shows the power-spectrum pre-factor $p(\ell)$. } 
    %
    \begin{tabular}{|l|l|c|c|c|p{0.21\textwidth}}
  \hline
  Case & ID & equation & $p(\ell)$ & $\nu(\ell)$ & Comment \\ \hline
  % 
  $1^{\rm st}$-order standard Limber, flat sky & L1Fl & (\ref{eq:P_ell_limber1}) + (i)
    & $1$ & $\ell$ & Pre-2014 CFHTLenS, DES-SV and DLS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, flat sky & ExtL1Fl & (\ref{eq:P_ell_limber1}) + (ii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ & $\ell + \frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $1^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL1FlHyb & (\ref{eq:P_ell_limber1}) + (iii)
    & $1$ & $\ell + \frac 1 2$ & Post-2014 CFHTLenS, KiDS \\ \hline
  %
  $1^{\rm st}$-order extended Limber, spherical & ExtL1Sph & (\ref{eq:C_ell_limber1})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+ \frac 1 2$ & \\ \hline
  %
  $2^{\rm st}$-order extended Limber, flat sky & ExtL2Fl &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $\frac{\ell^4}{(\ell+1/2)^4}$ 
    & $\ell+\frac 1 2$ & Converges only with ${\cal O}(\ell^{-1})$ \\ \hline
  %
  $2^{\rm st}$-order extended Limber, hybrid, flat sky & ExtL2FlHyb &  (\ref{eq:P_ell_limber1}) + (\ref{eq:P_ell_limber2}) + (iii)
    & $1$ 
    & $\ell+ \frac 1 2$ & Best flat-sky approximation \\ \hline
  %
  $2^{\rm st}$-order extended Limber, spherical & ExtL2Sph & (\ref{eq:C_ell_limber1})+(\ref{eq:C_ell_limber2_dr_s})
    & $\frac{\ellbar^2(\ell, 2)}{(\ell+1/2)^4}$ & $\ell+\frac 1 2$ & Best approximation \\ \hline
  %
  Full spherical & FullSph & (\ref{eq:C_ell_full}) &
      - & - & Correct projection \\ \hline
  %
  \end{tabular}

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by the spherical transform \citep{1973ApJ...185..413P}
%
\begin{equation}
 \xi_+(\theta) = \left\langle \gamma \gamma^\ast \right\rangle(\theta)
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$. This equation is easily derived
using the harmonics expansion of the shear (\ref{eq:gamma_harm_exp}) and the orthogonality relation of
the spherical harmonics (\ref{eq:Yellm_ortho}).

@@

We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to reduce computation time.
%
We find that the Hankel transform (equation~\ref{eqn:xiGG}) is accurate to better than $2$ percent level.
%
The difference between second-order Limber and full projection using the
spherical transform (\ref{eq:xi_Legendre}) is well below one percent for
$\vartheta < 200$ arcminutes. The red circles in Fig.~\ref{fig:xi_p_legendre}
present the limit of precision that we can achieve with the current fast Hankel
transform implementation for the correlation function. The green line shows the
limit of the second-order Limber approximation on the correlation function.

@@

  \caption{\emph{Left panel:} The shear correlation function component $\xi_+$
for the ExtL2Sph projection using the spherical transform (green line for
ExtL2Sph, grey squares for FullSph), compared to the flat-sky Hankel transform
of ExtL2Sph (red circles). \emph{Right panel:} The relative difference compared
to the full spherical case FullSph of ExtL2Sph with spherical transform (green
line), and Hankel transform (red circles).}

@@

To obtain the power spectrum of the lensing potential, 
we insert the lensing projection (equation~\ref{eq:psi}) into the
inverse harmonics expansion (equation~\ref{eq:psi_harm_exp_inv}) and write the 3D potential
as its Fourier transform (equation~\ref{eq:hatPhi_inv}) to get
%
\begin{equation}
  \psi_{\ell m} = \frac 2 {c^2} \int {\rm d} \Omega {\rm Y}^\ast_{\ell m}(\theta, \varphi)
    \int_0^\infty \frac{{\rm d}\chi}{\chi} q(\chi) \int \frac{{\rm d}^3 k}{(2\pi)^3} \hat \Phi(\vec k; \chi) {\rm e}^{-{\rm i} \vec k \cdot \vec r}.
\end{equation}
%
The 3D position vector $\vec r$ is a 3D position vector with polar coordinate
$r = \chi$ and polar angles $(\theta, \varphi)$. Similarly we denote with
$\theta_k, \varphi_k$ the polar angles of the 3D Fourier vector $\vec k$. We
insert the expansion of a plane wave into spherical harmonics,
%
% BBRV12 (41)
%
\begin{equation}
  {\rm e}^{{\rm i} \vec k \cdot \vec r} = 4 \pi \sum_{\ell=0}^{\infty} \sum_{m=-\ell}^{\ell}
    {\rm i}^\ell \, {\rm j}_\ell(k \chi)
    {\rm Y}_{\ell m}(\theta, \varphi) {\rm Y}_{\ell m}(\theta_k, \varphi_k) .
  \label{eq:wave_exp}
\end{equation}
%
Making use of the orthogonality of the spherical harmonics
%
% Rijt (6.15), Castro (B2).
%
\begin{equation}
  \int {\rm d} \Omega {\rm Y}^{}_{\ell m}(\theta, \varphi) {\rm Y}^\ast_{\ell^\prime m^\prime}(\theta, \varphi) = \delta_{\ell \ell^\prime} \delta_{m m^\prime},
  \label{eq:Yellm_ortho}
\end{equation}
%
the expression for $\psi_{\ell m}$ simplifies to
%
\begin{equation}
  \psi_{\ell m} = \frac {{\rm i}^\ell}{c^2 \pi^2} \int_0^\infty \frac{{\rm d}\chi}{\chi} q(\chi) \int {\rm d}^3 k \,
    \hat \Phi(\vec k; \chi) {\rm j}_\ell(k \chi) {\rm Y}_{\ell m}(\theta_k, \varphi_k).
  \label{eq:psi_ellm}
\end{equation}
%
To obtain the potential power spectrum, we take the absolute square of the
last equation and use the definition of the 3D potential power spectrum
(equation~\ref{eq:p_phi}). The delta-function resolves one 3D Fourier integral.
We split the second integration into radial and spherical coordinates, ${\rm
d}^3 k = {\rm d} k k^2 {\rm d} \Omega_k$ and use once again the orthogonality
of the spherical harmonics to resolve the spherical integral. This leads to the
potential power spectrum in equation~(\ref{eq:C_ell_phi_Pphi}).

@@

\cite{2008PhRvD..78d3002S} derive the lensing power spectrum in the flat-sky
limit, see their Equation (9). Inserting the Poisson equation and the
growth function, and writing the redshift filter function
$W_\kappa$ (eq.~10) in terms of the lensing efficiency and comoving distances,
$W_\kappa[z(\chi)] = H(z)^{-1} \chi q(\chi)$, their expression, using our
notations, reads
%
\begin{align}
  P_{ij}^\gamma(\ell) = & \frac 2 \pi \, \pref^2
                 \int_{0}^\infty {\rm d} \chi \chi \, \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^\infty {\rm d} \chi^\prime\, \chi^\prime
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty {\rm d} k \, k^2 \, P_{\rm m}(k; \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) ,
  \label{eq:P_ell_gamma}
\end{align}
%
This is consistent with our equation (\ref{eq:P_ell_gamma_full_der}) under the
additional assumption that mainly modes with $k \chi \approx k \chi^\prime
\approx \ell$ contribute to the integral; that is modes around the maxima of
the Bessel functions. Then, we can draw out of the integral the factor $\ell^4
\approx k^4 \chi^2 {\chi^\prime}^2$, to recover
equation~(\ref{eq:P_ell_gamma_full_der}).
In fact, using the approximation $k \chi = \ell$ seems to go too far, this is
already halfway the Limber approximation.
